### 2020-04-14 cmc-rnaseq-de/MSSM-Penn-Pitt_ACC_DLPFC_FEATURECOUNTS_GRCh38_CQN_Sparse.Rmd
note random effects

---
title: "MSSM-Penn-Pitt - ACC and DLPFC - DREAM - FEATURECOUNTS - GRCh38 - CQN - Model Selection"
author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
date: "`r date()`"
output: html_document
---

```{r knit2synapse, eval=FALSE}
library(synapser)
library(knit2synapse)

synLogin()

knit2synapse::createAndKnitToFolderEntity(file = "./cmc-rnaseq-de/MSSM-Penn-Pitt_ACC_DLPFC_FEATURECOUNTS_GRCh38_CQN_Sparse.Rmd",
                                 parentId ="syn21435176",
                                 folderName = "test mvBIC - MSSM-Penn-Pitt - ACC and DLPFC - DREAM - FEATURECOUNTS - GRCh38 - CQN")
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## Load required libraries
library(CovariateAnalysis) # get the package from devtools::install_github('th1vairam/CovariateAnalysis@dev')
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)
library(R.utils)
library(ggplot2)
library(limma)
library(psych)
library(edgeR)
library(cqn)
library(lme4)
library(biomaRt)
library(variancePartition)
library(synapser)
library(knitr)
library(githubr) # get the package from devtools::install_github('brian-bot/githubr')

library(mvBIC) # devtools::install_github("GabrielHoffman/mvBIC")

synLogin()

library(doParallel)
library(foreach)
library(future)

cl = makeCluster((availableCores()-2)/2)

registerDoParallel(cl)

options(xtable.type="html")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE, cache=TRUE}
parentId = 'syn21435176';
activityName = 'Covariate adjustments';
activityDescription = 'Covariate analysis of GRCh38 feature counts with CQN normalisation (ACC and DLPFC)';

thisFileName <- 'MSSM-Penn-Pitt_ACC_DLPFC_FEATURECOUNTS_GRCh38_CQN_Sparse.Rmd'

# Github link
thisRepo <- getRepo(repository = "kelshmo/analysis", ref="branch", refName='test-mvBIC')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('cmc-rnaseq-de/',thisFileName))
```
### Data download
#### Obtain count matrix and metadata from synapse.
```{r download.data, cache=TRUE}
downloadFile <- function(id){
  fread(synGet(id)$path, data.table = F)
}
downloadFile_version <- function(id , version){
  fread(synGet(id, version = version)$path, data.table = F)
}
# Download reprocessed counts
counts = 'syn21867938'
ALL_USED_IDs = counts
counts = downloadFile(counts) %>% data.frame()
counts$Chr <- NULL
counts$Start <- NULL
counts$End <- NULL
counts$Strand <- NULL

# Parse gene lengths
gene_lengths = counts[,c("Geneid", "Length")]
counts$Length <- NULL
gene_lengths <- dplyr::rename(gene_lengths, gene_id = Geneid)

# Get ancestry vector calculated using gemtools
ANCESTRY_ID = 'syn2511399'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = ANCESTRY_ID
ANCESTRY = downloadFile(ANCESTRY_ID) %>% 
  plyr::rename(c('DNA_report..Genotyping.Sample_ID' = 'SNP_report:Genotyping_Sample_ID'))

# Get genotype ids from synapse
GENOTYPE_ID = 'syn16816490'
ALL_USED_IDs = c(ALL_USED_IDs, GENOTYPE_ID)
GENOTYPE = downloadFile(GENOTYPE_ID) %>% 
  dplyr::select(Individual_ID, `SNP_report:Genotyping_Sample_ID`) %>%
  dplyr::inner_join(ANCESTRY)

#Get Metadata
metadata = 'syn16816488' 
ALL_USED_IDs[length(ALL_USED_IDs)+1] = metadata
metadata = downloadFile_version(metadata, version = 17)  %>% 
  dplyr::select(Individual_ID, Institution, Reported_Gender, Sex, 
                Ethnicity, ageOfDeath, `PMI_(in_hours)`, Dx,
                Sample_RNA_ID, one_of('rnaSeq_isolation:RIN',
                                      'rnaSeq_report:Ribozero_Batch', 'rnaSeq_report:Library_Batch', 
                                      'rnaSeq_report:Flowcell_Batch','rnaSeq_dissection:Brain_Region',
                                      'rnaSeq_report:Exclude?',
                                      'rnaSeq_report:Mapped_Reads',
                                      'rnaSeq_report:Intragenic_Rate','rnaSeq_report:Intronic_Rate',
                                      'rnaSeq_report:Intergenic_Rate','rnaSeq_report:Genes_Detected',
                                      'rnaSeq_report:Expression_Profiling_Efficiency',
                                      'rnaSeq_report:rRNA_Rate',
                                      'rnaSeq_report:Total_Reads', 'rnaSeq_report:Percent_Aligned'))

METADATA = metadata %>%
  dplyr::left_join(GENOTYPE) %>%
  dplyr::rename(PMI = `PMI_(in_hours)`,
                ReportExclude = `rnaSeq_report:Exclude?`,
                SampleID = Sample_RNA_ID,
                LibraryBatch = `rnaSeq_report:Library_Batch`,
                FlowcellBatch = `rnaSeq_report:Flowcell_Batch`,
                RibozeroBatch = `rnaSeq_report:Ribozero_Batch`,
                Tissue = `rnaSeq_dissection:Brain_Region`,
                MappedReads = `rnaSeq_report:Mapped_Reads`,
                IntragenicRate = `rnaSeq_report:Intragenic_Rate`, 
                IntronicRate = `rnaSeq_report:Intronic_Rate`, 
                IntergenicRate = `rnaSeq_report:Intergenic_Rate`, 
                GenesDetected = `rnaSeq_report:Genes_Detected`,
                ExpProfEfficiency = `rnaSeq_report:Expression_Profiling_Efficiency`, 
                rRNARate = `rnaSeq_report:rRNA_Rate`,
                TotalReads = `rnaSeq_report:Total_Reads`, 
                AlignmentRate = `rnaSeq_report:Percent_Aligned`,
                RIN = `rnaSeq_isolation:RIN`) %>%
  dplyr::select(SampleID, Individual_ID, Institution, Reported_Gender,Sex, Ethnicity, ageOfDeath, PMI, Dx, 
                RIN, ReportExclude, EV.1, EV.2, EV.3, EV.4, EV.5,
                LibraryBatch, FlowcellBatch, RibozeroBatch, Tissue, MappedReads, TotalReads, GenesDetected, AlignmentRate,
                IntragenicRate, IntergenicRate, IntronicRate, ExpProfEfficiency, rRNARate) %>% 
  dplyr::filter(Institution %in% c("MSSM", "Penn", "Pitt")) %>% 
  dplyr::filter(SampleID %in% colnames(counts), !is.na(SampleID)) %>%
  dplyr::mutate(Dx = forcats::fct_recode(Dx,  Other = "AFF", Other = "BP", Control = "Control", SCZ = "SCZ"))
```

### Data preprocessing
```{r preprocess.data}
ind = METADATA$SampleID[which(METADATA$ReportExclude == 1)]
writeLines(paste('Following',length(ind),'samples are marked exclude'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$Ethnicity) | is.na(METADATA$Institution) | is.na(METADATA$Dx)]
writeLines(paste('Following', length(ind), 'counts are missing clinical metadata'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$PMI)]
writeLines(paste('Following', length(ind), 'counts are missing PMI'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$Reported_Gender)]
writeLines(paste('Following', length(ind), 'counts are missing gender'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$ageOfDeath)]
writeLines(paste('Following', length(ind), 'counts are missing age of death'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$EV.1)]
writeLines(paste('Following', length(ind), 'counts are missing ancestry information'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))
```

```{r preprocess.data1, results='asis'}
# Match covariates to expression data
indToRetain = intersect(METADATA$SampleID, colnames(counts))

rownames(counts) <- counts$Geneid
counts = counts[,indToRetain]

rownames(METADATA) = METADATA$SampleID
METADATA = METADATA[indToRetain,]

METADATA %>% 
  group_by(Dx, Tissue) %>% 
  dplyr::summarise(count = n()) %>% 
  spread(Tissue, count) %>%
  kable()

METADATA = METADATA %>%
  dplyr::mutate(Dx.Tissue = paste(Dx, Tissue, sep = '.'))
```

```{r gene.param}
# Get GC content from biomart
backgroundGenes = data.frame(gene_id = rownames(counts)) %>%
  dplyr::mutate(id = gene_id) %>%
  tidyr::separate(id, c('ensembl_gene_id','position'), sep = '\\.')

# # Define biomart object
# mart <- useMart(biomart = "ENSEMBL_MART_ENSEMBL", 
#                 host = "uswest.ensembl.org", # Ensembl Release 99 (January 2020)
#                 dataset = "hsapiens_gene_ensembl")
# 
# # Query biomart
# Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol", "percentage_gene_gc_content", "gene_biotype", "chromosome_name"),
#                        filters = "ensembl_gene_id", values = backgroundGenes$ensembl_gene_id,
#                        mart = mart)
Ensemble2HGNC = downloadFile("syn21907981")
GENE.GC.CONT = Ensemble2HGNC %>%
  dplyr::left_join(backgroundGenes) %>%
  dplyr::select(gene_id, percentage_gene_gc_content, chromosome_name) %>%
  unique
rownames(GENE.GC.CONT) = GENE.GC.CONT$gene_id
```

<!-- ### Sex Check -->
<!-- ```{r Reported_Gender_plots} -->
<!-- counts$gene_id = rownames(counts)  -->
<!-- REPORTED.GENDER.COUNTS = GENE.GC.CONT %>%  -->
<!--   left_join(counts) %>% -->
<!--   dplyr::select(-one_of("percentage_gene_gc_content")) %>% -->
<!--   filter(chromosome_name == "X" |chromosome_name == "Y") %>%  -->
<!--   tidyr::gather(key = item, value = value, -c(gene_id, chromosome_name)) %>% -->
<!--   dplyr::mutate(value = log(value)) %>% -->
<!--   dplyr::rename(`counts(log)`= value) %>%  -->
<!--   dplyr::rename(SampleID = item) %>% -->
<!--   left_join(METADATA[,c("SampleID", "Reported_Gender")]) %>%  -->
<!--   dplyr::rename(`Reported Gender` = Reported_Gender)  -->

<!-- ##XIST and UTY expression  -->
<!-- #ENSG00000229807.11 and ENSG00000183878.15  -->
<!-- FILT <- REPORTED.GENDER.COUNTS %>%  -->
<!--   filter(gene_id == "ENSG00000229807.11" | gene_id == "ENSG00000183878.15") %>%  -->
<!--   dplyr::select(-one_of("chromosome_name")) %>%  -->
<!--   tidyr::spread(key = gene_id, value = `counts(log)`) %>%  -->
<!--   mutate(XIST = as.numeric(`ENSG00000229807.11`)) %>%  -->
<!--   mutate(UTY = as.numeric(`ENSG00000183878.15`)) -->

<!-- p = ggplot(FILT, aes (x= XIST, y = UTY))  -->
<!-- p = p + geom_point(aes(color=`Reported Gender`))  -->
<!-- p -->
<!-- ``` -->
### Covariates clustering
Determine relationship between covariates. 
```{r covariates.clustering}
FactorCovariates <- c('Individual_ID', "Institution", "Reported_Gender", "LibraryBatch", "Dx.Tissue", "RibozeroBatch", "FlowcellBatch")
ContCovariates <- c("ageOfDeath", "PMI", "RIN", "MappedReads", "IntragenicRate", "IntronicRate", "IntergenicRate",
                    "GenesDetected", "ExpProfEfficiency", "rRNARate", "TotalReads","AlignmentRate", 
                     "EV.1", "EV.2", "EV.3", "EV.4", "EV.5")

# Find inter relation between factor covariates
COVARIATES = METADATA[,c(FactorCovariates,ContCovariates),drop=F]
COVARIATES[,FactorCovariates] <- data.frame(lapply(COVARIATES[,FactorCovariates],function(x){
  x <- sapply(x,function(y){str_replace_all(as.character(y),'[^[:alnum:]]','_')})}))
rownames(COVARIATES) <- METADATA$SampleID

# Convert factor covariates to factors
COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], function(x){
  x = as.numeric(as.character(gsub('[\\,\\%]','',x)))
})

# Add in RIN^2 values
COVARIATES$RIN2 = COVARIATES$RIN^2
ContCovariates = c(ContCovariates, 'RIN2')
```
<!-- Correlation/association between covariates at an FDR <= 0.1 -->
<!-- ```{r covariates.correlation, fig.width=10, fig.height=8} -->
<!-- COVARIATES.CORRELATION = getAssociationStatistics(COVARIATES, PVAL = 0.05) -->
<!-- tmp = COVARIATES.CORRELATION$ESTIMATE -->
<!-- tmp[COVARIATES.CORRELATION$PVAL > 0.05] = 0 -->
<!-- h = Heatmap(tmp, col = colorRamp2(c(-1,0,1), c('blue','white','red')), name = 'AssocEstimate') -->
<!-- ComplexHeatmap::draw(h, heatmap_legend_side = 'left') -->
<!-- ``` -->

<!-- ## Explore metatdata -->
<!-- ```{r data.explore, fig.width = 12, fig.height = 12} -->
<!-- my.theme <- theme_bw() %+replace% theme(legend.position = 'top', axis.text.x = element_text(angle = 90, hjust = 1), plot.title=element_text(hjust=0.5)) -->

<!-- # RIN -->
<!-- p = list() -->
<!-- p[[1]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = RIN)) + geom_boxplot() -->
<!-- p[[1]] = p[[1]] + ggtitle('RIN') + my.theme -->

<!-- # Age of Death -->
<!-- p[[2]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = ageOfDeath)) + geom_boxplot() -->
<!-- p[[2]] = p[[2]] + ggtitle('AgeOfDeath') + my.theme -->

<!-- # PMI -->
<!-- p[[3]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = PMI)) + geom_boxplot() -->
<!-- p[[3]] = p[[3]] + ggtitle('PMI (in hours)') + my.theme -->

<!-- # Intronic Rate -->
<!-- p[[4]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = IntronicRate)) + geom_boxplot() -->
<!-- p[[4]] = p[[4]] + ggtitle('Intronic Rate') + my.theme -->

<!-- # IntergenicRate -->
<!-- p[[5]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = IntergenicRate)) + geom_boxplot() -->
<!-- p[[5]] = p[[5]] + ggtitle('Intergenic Rate') + my.theme -->

<!-- # Transcripts Detected -->
<!-- p[[6]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = IntragenicRate)) + geom_boxplot() -->
<!-- p[[6]] = p[[6]] + ggtitle('Intragenic Rate') + my.theme -->

<!-- # Mapped Reads -->
<!-- p[[7]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = MappedReads)) + geom_boxplot() -->
<!-- p[[7]] = p[[7]] + ggtitle('Mapped Reads') + my.theme -->

<!-- # PercentAligned -->
<!-- p[[8]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = TotalReads)) + geom_boxplot() -->
<!-- p[[8]] = p[[8]] + ggtitle('Total Reads') + my.theme -->

<!-- # rRNARate -->
<!-- p[[9]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = rRNARate)) + geom_boxplot() -->
<!-- p[[9]] = p[[9]] + ggtitle('rRNARate') + my.theme -->

<!-- multiplot(plotlist = p, cols = 3) -->

<!-- # Institution -->
<!-- # vcd::mosaic(~ Institution + Dx.Tissue, data = COVARIATES) -->

<!-- # Gender -->
<!-- # vcd::mosaic(~ Gender + Dx.Tissue, data = COVARIATES) -->
<!-- ``` -->
### Filter genes
Remove genes that have less than 1 cpm counts in at least 50% of samples per Dx.Tissue. Also remove genes with missing gene length and percentage GC content
```{r cpmnormalisation}
counts$gene_id = NULL

genesToAnalyze = dlply(METADATA, .(Dx.Tissue, Reported_Gender), .fun = function(mtd, count){
  processed.counts = getGeneFilteredGeneExprMatrix(count[,mtd$SampleID],
                                                   MIN_GENE_CPM=1, 
                                                   MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0.5)
  processed.counts$filteredExprMatrix$genes
}, counts)

genesToAnalyze = unlist(genesToAnalyze) %>% 
  unique() %>%
  intersect(GENE.GC.CONT$gene_id[!is.na(GENE.GC.CONT$percentage_gene_gc_content)]) %>%
  intersect(gene_lengths$gene_id[!is.na(gene_lengths$Length)])

#Subset union of all genes and convert subset to logcpm values 
PROCESSED_COUNTS = getGeneFilteredGeneExprMatrix(counts[genesToAnalyze, ], MIN_GENE_CPM=0, MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0)
pct.pc = PROCESSED_COUNTS$filteredExprMatrix$genes %>%
  dplyr::rename(gene_id = genes) %>%
  left_join(backgroundGenes) %>%
  left_join(Ensemble2HGNC) %>%
  group_by(gene_biotype) %>%
  dplyr::summarise(fraction  = n()) %>%
  filter(fraction > 100) %>%
  dplyr::mutate(fraction = fraction/length(PROCESSED_COUNTS$filteredExprMatrix$genes[,1]))
```
`r dim(PROCESSED_COUNTS$filteredExprMatrix$genes)[1]` genes are used for the analysis.

Following is the distribution of biotypes
`r kable(pct.pc)`

### Outlier analysis
Detect outlier samples based on their expression (logCPM) pattern
```{r detect.outliers, fig.height=8, fig.width=8, results='asis', cache = FALSE}
# Find principal components of expression to plot
PC <- prcomp(voom(PROCESSED_COUNTS$filteredExprMatrix$counts)$E, scale.=T, center = T)

# Plot first 2 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])

# Percentage from each PC
eigen <- PC$sdev^2
pc1 <- eigen[1]/sum(eigen)
pc2 <- eigen[2]/sum(eigen)

# Identify outliers - samples 4SDs from the mean
outliers <- as.character(plotdata$SampleID[c(which(plotdata$PC1 < mean(plotdata$PC1) - 4*sd(plotdata$PC1)),
                              which(plotdata$PC1 > mean(plotdata$PC1) + 4*sd(plotdata$PC1))), drop = T])

outliers <- c(outliers, as.character(plotdata$SampleID[c(which(plotdata$PC2 < mean(plotdata$PC2) - 4*sd(plotdata$PC2)),
                                           which(plotdata$PC2 > mean(plotdata$PC2) + 4*sd(plotdata$PC2))), drop = T] ))

plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, "SampleID")) %>%
    tidyr::separate(Dx.Tissue, c('Dx','Tissue'), sep = '_') %>%
  dplyr::mutate(label = SampleID) %>% 
  dplyr::mutate(label = ifelse((label %in% outliers), label, NA))

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=Institution, shape=Tissue, size=ageOfDeath))
p <- p + theme_bw() + theme(legend.position="right") + facet_grid(Tissue~.)
p <- p + geom_text(aes(label= label), size=4, hjust=0)
p

# Plot abberent distribution of logcpm counts
tmp1 = voom(PROCESSED_COUNTS$filteredExprMatrix$counts)$E %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  tidyr::gather(SampleID, logCPM, -ensembl_gene_id) %>%
  left_join(COVARIATES %>%
              rownameToFirstColumn('SampleID') %>%
              tidyr::separate(Dx.Tissue, c('Dx', 'Tissue'), sep = '_'))

p = ggplot(tmp1 %>%
             dplyr::filter(SampleID %in% outliers),
           aes(x = logCPM, color = SampleID)) + geom_density() 
p = p + theme(legend.position = 'top') + facet_grid(Dx+.~Tissue, scale = 'free')
p

indToRetain = setdiff(colnames(PROCESSED_COUNTS$filteredExprMatrix$counts), outliers)
PROCESSED_COUNTS$filteredExprMatrix$counts = PROCESSED_COUNTS$filteredExprMatrix$counts[,indToRetain]
COVARIATES = COVARIATES[indToRetain,]

tmp = tidyr::separate(COVARIATES, Dx.Tissue, c('Dx', 'Tissue'), sep = '_') %>%
  group_by(Dx, Tissue) %>%
  dplyr::summarise(count = n()) %>%
  spread(Tissue, count)

NEW.COUNTS = PROCESSED_COUNTS$filteredExprMatrix$counts
```
Processing `r dim(PROCESSED_COUNTS$filteredExprMatrix)[1]` genes in `r dim(PROCESSED_COUNTS$filteredExprMatrix)[2]` samples from `r length(unique(COVARIATES$Individual_ID))` unique individuals

Based on the expression pattern following samples were tagged as outliers: `r paste(outliers, collapse = ', ')`

Distribution of samples are:
`r kable(tmp)`

### Library Normalisation
Initial library normalisation usign cqn
```{r cqn}
# Compute offset for gene length and gc content
gene_lengths = tibble::column_to_rownames(gene_lengths, var = 'gene_id') %>% as.data.frame()
CQN.GENE_EXPRESSION = cqn::cqn(NEW.COUNTS, 
                          x = GENE.GC.CONT[PROCESSED_COUNTS$filteredExprMatrix$genes$genes, 'percentage_gene_gc_content'],
                          lengths = gene_lengths[PROCESSED_COUNTS$filteredExprMatrix$genes$genes, 'Length'],
                          lengthMethod = "smooth", 
                          verbose = FALSE)
CQN.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION$y + CQN.GENE_EXPRESSION$offset
```

<!-- ### Co-expression distribution -->
<!-- Coexpression of genes  -->
<!-- ```{r coexp1, cache=FALSE, fig.height=5, fig.width=5} -->
<!-- cr = cor(t(CQN.GENE_EXPRESSION$E)) -->
<!-- hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation') -->
<!-- ``` -->

<!-- ### Sample clustering -->
<!-- ```{r decompse.normalise.data, fig.height=8, fig.width=8, results='asis', cache=FALSE} -->
<!-- # Find principal components of expression to plot -->
<!-- PC <- prcomp(CQN.GENE_EXPRESSION$E, scale.=T, center = T) -->

<!-- # Plot first 2 PCs -->
<!-- plotdata <- data.frame(SampleID=rownames(PC$rotation),  -->
<!--                        PC1=PC$rotation[,1],  -->
<!--                        PC2=PC$rotation[,2]) -->

<!-- plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>% -->
<!--   tidyr::separate(Dx.Tissue, c('Dx','Tissue'), sep = '_') -->

<!-- p <- ggplot(plotdata, aes(x=PC1, y=PC2)) -->
<!-- p <- p + geom_point(aes(color=Institution, shape=Tissue, size=ageOfDeath)) -->
<!-- p <- p + theme_bw() + theme(legend.position="right") + facet_grid(Tissue~.) -->
<!-- # p <- p + geom_text(aes(label= SampleID), size=4, hjust=0) -->
<!-- p -->
<!-- ``` -->
<!-- Tree based clustering of samples -->
<!-- ```{r decompse.normalise.data.1, fig.height=6, fig.width=10, results='asis'} -->
<!-- # Eucledian tree based analysis -->
<!-- COVARIATES.tmp = data.matrix(COVARIATES[,c("Institution", "Reported_Gender", "LibraryBatch", "Dx.Tissue")]) -->
<!-- COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0 -->

<!-- tree = hclust(as.dist(t(CQN.GENE_EXPRESSION$E))) -->
<!-- cols = WGCNA::labels2colors(COVARIATES.tmp); -->

<!-- WGCNA::plotDendroAndColors(tree,  -->
<!--                            colors = cols,  -->
<!--                            dendroLabels = FALSE,  -->
<!--                            abHeight = 0.80,  -->
<!--                            main = "Sample dendrogram", -->
<!--                            groupLabels = colnames(COVARIATES.tmp)) -->
<!-- ``` -->
<!-- ```{r temp, include = F} -->
<!-- dev.off() -->
<!-- gc() -->
<!-- ``` -->

<!-- ### Distribution of samples (log cpm) -->
<!-- ```{r lcpm.dist, cache=FALSE} -->
<!-- # Plot abberent distribution of logcpm counts -->
<!-- tmp1 = (CQN.GENE_EXPRESSION$E) %>% -->
<!--   rownameToFirstColumn('gene_id') %>% -->
<!--   tidyr::gather(SampleID, logCPM, -gene_id) %>% -->
<!--   left_join(COVARIATES %>% -->
<!--               rownameToFirstColumn('SampleID')) -->

<!-- p = ggplot(tmp1, aes(x = logCPM, color = SampleID)) + geom_density()  -->
<!-- p = p + theme_bw() %+replace% theme(legend.position = 'NONE') + facet_grid(.~Dx.Tissue, scale = 'free') -->
<!-- p -->
<!-- ``` -->

### Subset Genes
```{r subset-genes}
rand_sample <- sample(1:dim(NEW.COUNTS)[1], size = round(dim(NEW.COUNTS)[1]*.6), replace = FALSE)
subset_NEW.COUNTS<- NEW.COUNTS[rand_sample,]

subset_CQN <- CQN.GENE_EXPRESSION$E[rand_sample,]
```

#### Phase I (clinical, ancestry and sample specific technical variables)
```{r phase1}
# Fit a base model with the following covariates
base_covars <- "Dx.Tissue"
random_effect <- "Individual_ID"

# Set base formula
base_formula <- glue::glue("~ ", glue::glue_collapse(base_covars, sep = " + ")) %>% 
  glue::glue(" + (1|{random_effect})")

# Stepwise regression
phase1 <- mvBIC::mvForwardStepwise(exprObj = subset_CQN,
                         baseFormula = base_formula, 
                         data = COVARIATES,
                         variables = array(c("(1|Institution)", "Reported_Gender", "PMI", 
                                                             "ageOfDeath", "RIN", "RIN2", "EV.1", 
                                                             "EV.2", "EV.3", "EV.4", "EV.5")))
phase1
```

#### Phase II (batch effects)
```{r phase2}
phase2 <- mvBIC::mvForwardStepwise(exprObj = subset_CQN,
                         baseFormula = phase1$formula, 
                         data = COVARIATES,
                         variables = array(c("(1|LibraryBatch)", "(1|FlowcellBatch)")))
phase2
```

#### Phase III (RNASeq alignment specific covariates)
```{r phase3}
phase3 <- mvBIC::mvForwardStepwise(exprObj = subset_CQN,
                         baseFormula = phase2$formula, 
                         data = COVARIATES,
                         variables = array(c("MappedReads", "IntragenicRate", "IntronicRate", "IntergenicRate", "GenesDetected", "rRNARate", "TotalReads")))
phase3
```

#### Variables included in the model 
`r phase3$formula`

<!-- ### Store files in synapse -->
<!-- ```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE} -->
<!-- # Set annotations -->
<!-- all.annotations = list( -->
<!--   dataType = 'geneExpression', -->
<!--   dataSubtype = 'processed', -->
<!--   assay	 = 'rnaSeq', -->

<!--   tissue	= 'dorsolateral prefrontal cortex, anterior cingulate cortex',  -->
<!--   study = 'CMC',  -->

<!--   species = 'Human', -->
<!--   consortium	= 'CMC', -->

<!--   analysisType	= "data normalization", -->

<!--   transcriptNormalizationMethod	= 'CQN', -->
<!--   transcriptQuantificationMethod = 'featureCounts', -->
<!--   referenceSet = 'GRCh38' -->
<!-- ) -->

# Code
``` {r save_code}
CODE <- Folder(name = "test mvBIC - MSSM-Penn-Pitt - ACC and DLPFC - DREAM - FEATURECOUNTS - GRCh38 - CQN", parentId = parentId)
CODE <- synStore(CODE)
```
<!-- # Store covariates -->
<!-- COVARIATES = rownameToFirstColumn(COVARIATES, 'SampleID') -->
<!-- write.table(COVARIATES, file = 'CMC_ACC_DLPFC_Covariates.tsv', sep = '\t', row.names=F, quote=F) -->
<!-- COV_OBJ = synapser::File('CMC_ACC_DLPFC_Covariates.tsv', name = 'Covariates', parentId = CODE$properties$id) -->
<!-- COV_OBJ = synStore(COV_OBJ, used = ALL_USED_IDs, activityName = activityName,  -->
<!--                       executed = thisFile, activityDescription = activityDescription) -->
<!-- synSetAnnotations(COV_OBJ, annotations = all.annotations) -->

<!-- # Store filtered counts -->
<!-- PROCESSED_COUNTS$filteredExprMatrix$counts %>% -->
<!--   rownameToFirstColumn('ensembl_gene_id') %>% -->
<!--   write.table(file = 'CMC_ACC_DLPFC_Counts.tsv', sep = '\t', row.names=F, quote=F) -->
<!-- COUNT_OBJ = File('CMC_ACC_DLPFC_Counts.tsv', name = 'Counts (filtered raw)', parentId = CODE$properties$id) -->
<!-- COUNT_OBJ = synStore(COUNT_OBJ, used = ALL_USED_IDs, activityName = activityName,  -->
<!--                       executed = thisFile, activityDescription = activityDescription) -->
<!-- synSetAnnotations(COUNT_OBJ, annotations = all.annotations) -->

<!-- # Store logCPM -->
<!-- CQN.GENE_EXPRESSION$E %>% -->
<!--   rownameToFirstColumn('ensembl_gene_id') %>% -->
<!--   write.table(file = 'CMC_ACC_DLPFC_logCPM.tsv', sep = '\t', row.names=F, quote=F) -->
<!-- LCOUNT_OBJ = File('CMC_ACC_DLPFC_logCPM.tsv', name = 'Counts (filtered logCPM)', parentId = CODE$properties$id) -->
<!-- LCOUNT_OBJ = synStore(LCOUNT_OBJ, used = ALL_USED_IDs, activityName = activityName,  -->
<!--                       executed = thisFile, activityDescription = activityDescription) -->
<!-- synSetAnnotations(LCOUNT_OBJ, annotations = all.annotations) -->

<!-- # Store cqn offsets -->
<!-- CQN.GENE_EXPRESSION$offset %>% -->
<!--   rownameToFirstColumn('ensembl_gene_id') %>% -->
<!--   write.table(file = 'CMC_ACC_DLPFC_offset.tsv', sep = '\t', row.names=F, quote=F) -->
<!-- OFFSET_OBJ = File('CMC_ACC_DLPFC_offset.tsv', name = 'Gene length and GC content offset', parentId = CODE$properties$id) -->
<!-- OFFSET_OBJ = synStore(OFFSET_OBJ, used = ALL_USED_IDs, activityName = activityName,  -->
<!--                       executed = thisFile, activityDescription = activityDescription) -->
<!-- synSetAnnotations(OFFSET_OBJ, annotations = all.annotations) -->
<!-- ``` -->

<!-- |  *Results*                                  |  *SynapseID*                     | -->
<!-- |  -----------------------------------------  |   ---------                      | -->
<!-- |  Covariates                                 |  `r COV_OBJ$properties$id`       | -->
<!-- |  Counts (raw)                               |  `r COUNT_OBJ$properties$id`     | -->
<!-- |  Counts (lcpm)                              |  `r LCOUNT_OBJ$properties$id`    | -->
<!-- |  Offset (for gene length and gc content)    |  `r OFFSET_OBJ$properties$id`    | -->

### Source Rmd
[Source markdown](`r thisFile`)